Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2020-11-03 Manuscript/manuscriptPGPEsuppl.tex
Update dihedral and ion binding results in the manuscript
Update manuscript

\usepackage[version=3]{mhchem}
\usepackage[T1]{fontenc}
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\usepackage{rotating}
\usepackage{upgreek}				
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xr}
\externaldocument{manuscriptPGPE}
\usepackage{soul} % for highlights with \hl{}

@@

@@

@@

\section{Calcium binding to POPC:POPG (4:1) mixtures}
\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC4.eps}
  \caption{\label{changesWITHCaClPG1PC4}
    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (4:1)
    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{macdonald87} and simulations.
    The changes with respect to the systems without CaCl$_2$ are shown for other data than
    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
  }
  \todo{CHARMM36 simulations should be longer and with Na couterions.}
\end{figure*}

>>>>

\usepackage[version=3]{mhchem}
\usepackage[T1]{fontenc}
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
%\newcommand{\todo}[1]{\textcolor{red}{#1}}
\usepackage[obeyFinal]{easy-todo}
\usepackage{rotating}
\usepackage{upgreek}				
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xr}
\externaldocument{manuscriptPGPE}
\usepackage{soul} % for highlights with \hl{}

@@

\section{Comparison of headgroup order parameters from different force fields to experiments}
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.pdf}
  \caption{\label{HGorderParametersPE}
    The headgroup and glycerol backbone order parameters of PE lipids
    from experiments (POPE and signs this work, DPPE from Ref.~\citenum{seelig76}
    and E.coliPE from Ref.~\citenum{gally81}) and simulations with different force fields.
  }
  \todo{This should be clarified as in NMRlipidsI and error bars should be added.
    Probably larger error bars for united atom models based on the report by Fuchs et al.
  } \\
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
  \caption{\label{HGorderParametersPOPG}
    The headgroup and glycerol backbone order parameters of PG lipids
    from experiments (POPG and signs from this work and from Ref.~\citenum{borle85}, %contains 10mM of PIPES,
    DPPG with 100mM NaCl from Ref.~\citenum{wohlgemuth80},% contains 10mM PIPES and,
    and E.Coli PG results from Ref.~\citenum{gally81}).
    and simulations with different force fields.
  }
\end{figure}

\pagebreak

\subsection{PC headgroup interactions with PE and PG}
\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
  \caption{\label{HGorderparametersPCvsPEPG}
    Modulation of POPC headgroup order parameters with increasing amount of POPE (left) and POPG (right) in bilayer
    from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{P-N vector angles from Berger-OPLS and GROMOS-CKP simulations are yet to be analyzed.}
\end{figure*}

In experiments, the PC headgroup order parameters increase
with the addition of negatively charged PG or PS lipids, but
are not affected by the addition of zwitterionic PE and SM lipids
or cholesterol (Fig.~\ref{HGorderparametersPCvsPEPG}).
This can be explained by the electrometer concept, which suggests that the 
headgroup dipole tilts more parallel to the membrane plane upon addition of
negative charge to the membrane~\cite{seelig87, scherer87,antila18}.
The response of PC headgroup order parameters to PE by the tested CHARMM36 and
Berger-OPLS force fields, although CHARMM36 slightly overestimates the changes (Fig.~\ref{HGorderparametersPCvsPEPG}).
The good performance of Berger-OPLS simulations is notable because the response of headgroup order parameters
to cholesterol was significantly overestimated by the Berger/Höltje force field in our previous work \cite{botan15}.
\todo{This is text by P. Fuchs, copied from the blog. \\
Area results in nm$^2$, the error is $<=$ 0.003 nm$^2$ \\
- pure POPC \\
CHARMM36: 0.624 \\
Berger : 0.649 \\
- POPC/POPE 50:50 \\
CHARMM36 : POPC 0.609, POPE 0.557 \\
Berger-hacked: POPC 0.637, POPE 0.632 \\
----- \\
One can see that CHARMM 36 predicts a drop in the area on going from pure POPC to POPC/POPE 50:50. This means that POPC pack tightly to POPE. 
In contrast, the values for Berger are not that changed. 
The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.\\
-------\\
The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
which supports the more condensed PE bilayer.
This is interesting, but not exactly the core message of the manuscript.
Maybe we should mention this very briefly? For example, we could just report the areas per lipid (without distinguishing PC and PE)
and mention the difference between CHARMM36 and Berger. I have opened an issue for this: https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/7
}

None of the force fields fully reproduces the PC headgroup order parameter response to the increasing amount of
PG, which may be related to the counterion binding affinity (see also the next section) \cite{antila19}.
In all force fields except Slipids, the order parameters of different
hydrogens attached to the $\alpha$-carbon are responsing differently when mixed with PE or PG lipids
\todo{Maybe we should figure out what is the reason for this? \\
  Maybe we should analyze the P-N vector angle from different simulations? \\
  https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/10}.

For $\beta$-carbon order parameter in PG headgroup, experiments report
mild increase \cite{macdonald87} or no change~\cite{borle85} upon addition 
of PC lipids (Fig. \ref{HGorderparametersPGvsPCchange}). 
Simulations with all the tested force fields give only very small changes also for
the $\alpha$-carbon order parameter (Figs. \ref{HGorderparametersPGvsPC} and \ref{HGorderparametersPGvsPCchange}). 
Therefore, the simulations are generally in line with experiments, suggesting that the
interactions with PC do not essentially effect the PG headgroup structure.
This suggests that the interactions between PG and PC headgroups are captured better in
simulations than for PS headgroup, where
all the force fieds significantly overestimated the structural response of PS headgroup
to the interactions with PC lipids \cite{antila18}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPCchange.eps}
  \caption{\label{HGorderparametersPGvsPCchange}
    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
  }
  \todo{P-N angles from Slipids and GROMOS-CKP yet to be calculated.}
\end{figure}

\clearpage
\subsection{Sodium binding to PE and PG lipid bilayers}
Sodium binding affinity to PE lipids has not been measured experimentally,
but large differences to PC would be surprising. 
In simulations, the sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
but lesser extend than reported previously for PC~\cite{catte16}.
\todo{This will be finished once we have all the simulation details and Lipid17 simulations with correct dihedrals from issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12},
Because some simulation and ion parameters are not identical with the previous work~\cite{catte16},
we compare POPE results to the POPC simulations ran with identical parameters (Fig. \ref{NAdensPC}).
In Lipid17 with the strongest sodium binding affinity to POPE,
the binding affinity is approximately similar to POPC.
Slipids and CHARMM36 exhibit slightly, and
GROMOS-CKP subtantially weaker binding to POPE than to POPC.
%The Lipid17 exhibits stronger sodium binding affinity
%to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
%Experimental data for sodium binding to PE lipids is not available, while
%simulations in agreement with electrometer data from NMR suggests that
%sodium binding to POPC is weaker than in any of the simulations
%here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}.
Assuming that the binding
to POPE would be similar than to POPC, the sodium binding affinity to POPE
is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
but substantially overestimated in Lipid17 simulation.
%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
%previous Amber lipid version, Lipid14.
%\todo{The potential differences in the used ion models should be checked}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
  \caption{\label{NAdensPE}
    Sodium (solid line) and choride ion density profiles along membrane normal
    from different simulations with PE lipids.
  }
\end{figure}

Simulations with PG lipids give similar dependence on force field as observed in POPE simulations:
Lipid 17 simulations with Dang ion parameters exhibits stronger counter-ion binding affinity to pure POPG bilayer
than CHARMM36, Slipids, and GROMOS-CKP simulations, which are roughly similar (Fig. \ref{CIdensPG}).
Lipid17 also exhibits less increase in POPC headgroup
%$\beta$ carbon
order parameters upon addition of POPG than other simulations (Fig. \ref{HGorderparametersPCvsPEPG}),
and lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$).
In our previous study about PS lipids \cite{antila18}, such behaviour was related to 
%the results suggest that
the overestimated counterions binding and shielding the electrostatic repulsion between PG headgroups
in bilayers.% simulated with Lipid17 parameters.
%the underestimated increase of PC
%headgroup order parameters upon addition of negaticely charged lipids
%was related
%to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
%Similar correlation is observed here in %other than GROMOS-CKP simulation:
%%Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
%
Even though the area per lipid in CHARMM36, Slipids, and GROMOS-CKP simulations is in good agreement
with experiments (Fig. \ref{CIdensPG}), the experimental increase in 
%In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of
POPC headgroup order parameters upon addition of POPG are not fully reproduced (Fig. \ref{HGorderparametersPCvsPEPG}).
%more than in CHARMM36 and Slipids simulations even though the binding affinities in these
%simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
%are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
%and other decreases \todo{Can we figure out the reason for this?}.
Therefore we conclude that the counter-ion binding affinity is overestimated in Lipid17 simulations,
% with the strongest affinity among the tested models,
while the other simulations are more realistic, but slight overbinding cannot be excluded.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities and area per lipids from simulations with PG lipids.
    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
  }
\end{figure}

@@

%\begin{table*}[htb]
%
\begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with PE lipids.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systemsPE}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      POPE  & CHARMM36 \cite{??}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
      POPE  & CHARMM36 \cite{??}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
      POPE  & CHARMM36 \cite{??}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
      POPE  & CHARMM36ua \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
      \hline
      DPPE  & Slipids \cite{jambeck12b}    &0    & 288 	& 9386  &0    & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
      POPE  & Slipids \cite{jambeck12b} &0    & 336	& ?     &0    & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
      POPE  & Slipids \cite{jambeck12b}            & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
      \hline
      DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3655  &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
      POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
      POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
      \hline
      POPE  & GROMOS 43A1-S3 \cite{??}     &0    & 128	& 3552     &0    & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles}  \\
      \hline
      POPE  & OPLS-UA vdW on H \cite{??}   &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
      POPE  & OPLS-UA \cite{??}            &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \\
      \hline
      POPE  & OPLS-MacRog \cite{rog16}     &0    & 144	& 5760     &0    & 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
      \hline
      POPE  & Berger-Vries \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
      POPE  & Berger-largeH \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
      DOPE  & Berger-Vries \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
      DOPE  & Berger-largeH \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
      \hline
      POPE             & LIPID17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
    \end{tabular}
  \end{minipage}
  \todo{Citation for CHARMM36 PE?} \\
  \todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
  \todo{Citation for GROMOS-CKP?} \\
  \todo{Citation for GROMOS 43A1-S3?} \\
  \todo{Citation for OPLS-UA models?} \\
  \todo{Citations for Berger-* simulations?} \\
  \todo{LIPID17 simulations with correct dihedrals still coming}
%\end{table*}
  \end{sidewaystable*}

  %    \begin{table*}[htb]
  \begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with PG lipids.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systemsPG}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}    &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
      POPG             & CHARMM36 \cite{??}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
      POPG             & CHARMM36 \cite{??}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
      \hline
      POPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 10664   &0     & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232  &0     & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232   &0     & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0    & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGslipids} \\
      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
      \hline
      POPG             & LIPID17 / Dang \cite{gould18,smith94}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
      POPG             & LIPID17 \cite{??}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
      \hline
      POPG             & GROMOS-CKP \cite{??}         & 0              & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP} \\
      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 49 &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
    \end{tabular}
  \end{minipage}
    \todo{Citations and ion model for CHARMM36?} \\
  \todo{Lipid17 simulation with correct dihedral potentials still coming.} \\
  \todo{Citation and ion model for GROMOS-CKP?}
%\end{table*}
 \end{sidewaystable*}

%\begin{table*}[htb]
  \begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with PE and PG lipids mixed with PC.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systemsMIX}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
%      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
%      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
      POPC                   & CHARMM36 \cite{??}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
      \hline
      POPC             & CHARMM36 \cite{??}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
      POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
      \hline
      POPC             & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPCmacrog300K} \\
      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1macrogT300K} \\
      \hline
      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  298  & 170 & 100 & \cite{POPCslipid298K} \\
      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120  & 0  &  298  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\
     \hline
      POPC                   & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCgromosCKPT310K}  \\
      POPC:POPG (7:3)        & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG3gromosCKPT310K} \\
     \hline
      POPC                   & Slipid \cite{jambeck12b}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
      POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
    \end{tabular}
  \end{minipage}
    \todo{Citation and ion model for GROMOS-CKP?} \\
  \todo{Citation and description for ''Berger'' model?} \\
  \todo{Lipid17 POPC and POPC:POPG mixtures (https://doi.org/10.5281/zenodo.3241242 and https://doi.org/10.5281/zenodo.3237656)
    should be added after simulated with corrected dihedrals.}\\
  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
%\end{table*}
\end{sidewaystable*}

  \begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with PE and PG lipids mixed with PC.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systemsMIX}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 150:150 & 31401 & 57 &  298  & 718 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
      \hline
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 150:150 & 29865 & 569 &  298  & 600 & 400 & \cite{Lipid17eccPOPCPOPG50501000mMCaCl} \\
      \hline
      POPC             & Berger \cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  &0  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
      POPC:POPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
      POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
     \hline
      DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
      DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}  &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
    \end{tabular}
  \end{minipage}
    \todo{Citation and ion model for GROMOS-CKP?} \\
  \todo{Citation and description for ''Berger'' model?} \\
  \todo{Lipid17 POPC and POPC:POPG mixtures (https://doi.org/10.5281/zenodo.3241242 and https://doi.org/10.5281/zenodo.3237656)
    should be added after simulated with corrected dihedrals.}\\
  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
%\end{table*}
\end{sidewaystable*}

@@

\section{Calcium binding to POPC:POPG (4:1) mixtures}
\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC4.eps}
  \caption{\label{changesWITHCaClPG1PC4}
    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (4:1)
    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{macdonald87} and simulations.
    The changes with respect to the systems without CaCl$_2$ are shown for other data than
    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
  }
  \todo{Lipid17ecc data to be analyzed and added.}
\end{figure*}
